Police Shootings and Racism in America

### 2020-10-17 police_eda.Rmd
cleaning office race

## Load Data
```{r load-data}
all <- read.csv("allshootingsdata.csv")

philly <- read.csv("philly_shootings.csv")
cincinnati <- read.csv("cincinnati_shootings.csv")
```

```{r clean-all}
all <- all %>% 
  select(-c("Date", "NatureOfStop", "Notes", "SubjectAge", "NumberOfShots", "Department", "FullNarrative", "fdate", "month", "day", "b", "l", "w", "a", "tr", "AvgAge", "VA", "F", "nshots", "sit"))

@@

@@

```{r clean-cities}
cinn <- cincinnati %>% 
  select(RACE, SEX, TYPE) %>% 
  rename(race = RACE, gender = SEX, fatal = TYPE) %>% 
  mutate(city = "Cincinnati")

phil <- philly %>% 
  select(race, sex, fatal) %>% 
  rename(gender = sex) %>% 
  mutate(city = "Philadelphia")

comp <- rbind(cinn, phil)

comp <- comp %>%
  filter(race != "UNKNOWN") %>% 
  filter (race != "") %>% 
  filter(race != "UNK") %>% 
  filter(gender != "")

comp$race[comp$race == "WHITE"] <- "W"
comp$race[comp$race == "BLACK"] <- "B"
comp$race[comp$race == "OTHER"] <- "O"
comp$race[comp$race == "HISPANIC"] <- "H"

comp$gender[comp$gender == "MALE"] <- "M"
comp$gender[comp$gender == "FEMALE"] <- "F"

comp$fatal[comp$fatal == "0"] <- "N"
comp$fatal[comp$fatal == "1"] <- "F"
comp$fatal[comp$fatal == "NONFATAL"] <- "N"
comp$fatal[comp$fatal == "FATAL"] <- "F"
```

```{r fatal-eda}
temp <- comp %>% 
  group_by(fatal) %>% 
  summarise(count = n())

ggplot(temp, aes(x=fatal, y=count)) + 
  geom_bar(stat="identity") + 
  labs(x = "Fatal", y = "Count", title = "Counts for Fatal vs. Non-Fatal Police Shootings") 
```
```{r race-eda-cities}
temp <- comp %>% 
  group_by(fatal, race) %>% 
  summarise(count = n())

ggplot(temp, aes(x=race, y=count, fill = fatal)) + 
  geom_bar(position="dodge", stat="identity") + 
  labs(x = "Race", y = "Count", title = "Counts for Victim Race by Fatal vs. Non-Fatal in All Shootings") 
```

```{r gender-eda-cities}
temp <- comp %>% 
  group_by(fatal, gender) %>% 
  summarise(count = n())

ggplot(temp, aes(x=gender, y=count, fill = fatal)) + 
  geom_bar(position="dodge", stat="identity") + 
  labs(x = "Gender", y = "Count", title = "Counts for Victim Gender by Fatal vs. Non-Fatal in All Shootings") 
```

>>>>

## Load Data

```{r load-data}
# Shootings
all <- read.csv("data/allshootingsdata.csv")

# Census Predictors
census <- read.csv("data/2013census.csv")[,-c(8,9)]
states <- lapply(census$department, function(x) str_extract_all(x,"\\([^()]+\\)")[[1]])
states <- substring(states, 2, nchar(states)-1)
census$state <- states

# Gun Law Predictors
gun_provisions <- read.csv("data/gun_provisions.csv") %>% 
  filter(year == 2013)
gun_provisions <- gun_provisions %>% 
  mutate(total_pr = rowSums(gun_provisions[,3:136]),
         deesc_train = ifelse(state %in% c("California", "New Mexico", "Texas", "Oklahoma", "South Dakota", "Illinois", "Indiana", "Tennessee", "Washington", "Missouri", "Georgia", "Ohio", "Maryland", "Massachusetts", "Connecticut", "Maine"), 1, 0)) %>% 
  select(state, total_pr, deesc_train)
gun_provisions$state <- state.abb[match(gun_provisions$state,state.name)]
```

```{r clean-all}
all <- all %>% 
  select(-c("Date", "NatureOfStop", "Notes", "SubjectAge", "NumberOfShots", "Department", "FullNarrative", "fdate", "month", "day", "tr", "AvgAge", "VA", "F", "nshots", "sit")) %>% 
  mutate(other = as.numeric(other), 
         replica = as.numeric(replica), 
         knife = as.numeric(knife), 
         gun = as.numeric(weapon == "gun"), 
         unarmed = as.numeric(weapon == "unarmed"),
         unknown_weapon = as.numeric(weapon == "U"),
         id = paste0("C", row_number())) %>% 
  select(-weapon)

@@

```{r split_data}
single_victims <- all %>% 
  filter(!str_detect(SubjectRace, ";"))

mult_victims <- all %>% 
  filter(str_detect(SubjectRace, ";"))
```

```{r}
temp <- single_victims %>% 
  group_by(Fatal, SubjectRace) %>% 
  summarise(count = n())

ggplot(single_victims, aes(x = Fatal, y = count)) +
  geom_bar(stat="identity") + 
  labs(x = "Fatal", y = "Count", title = "Counts for Fatal vs. Non-Fatal Police Shootings") 
  group_by(Fatal, race) %>% 
  summarise(count = n())
```

@@

### 2020-10-17 submissionii.Rmd
cleaning office race

```{r clean-all}
all <- all %>% 
  select(-c("Date", "NatureOfStop", "Notes", "SubjectAge", "NumberOfShots", "Department", "FullNarrative", "fdate", "month", "day", "tr", "AvgAge", "VA", "F", "nshots", "sit")) %>% 
  mutate(other = as.numeric(other), 
         replica = as.numeric(replica), 
         knife = as.numeric(knife), 
         gun = as.numeric(weapon == "gun"), 
         unarmed = as.numeric(weapon == "unarmed"),
         unknown_weapon = as.numeric(weapon == "U"),
         id = paste0("C", row_number())) %>% 
  select(-weapon)

@@

|        As a result, I have decided to build off of VICE News’ study and investigate data on police shootings further to understand the roles that the race of both victim and officer, as well as other factors such as their genders and whether they are carrying a weapon, play in fatal versus non-fatal police shootings. Additionally, to account for the varying locations of the homicides, I will be using a dataset found on Kaggle detailing the gun provisions that are upheld by each state see how gun legislation affects lethal vs. non-lethal shootings. I will also add a predictor indicating whether the state in which the homicide occurred requires de-escalation training for police officers. Finally, I will explore how the ratio of black to white population in each location relates to police use of deadly force. This will allow me to better understand how racism has manifested itself in America’s police system and determine whether current attempts to prevent police use of lethal force are effective or not.

@@

|       Finally, to examine city demographics as a predictor, I will be using population data from the 2013 American Community Survey that VICE News had already cleaned and standardized to the shootings dataset, meaning the city names can be matched up between the two datasets. VICE News most likely provided ony 2013 census data because it is the average year of all the years represented in the shootings dataset. The census data includes 7 variables: the police department and the city it's located in, as well as the city's black, asian, hispanic, white, and overall total population in 2013.

@@

>>>>

```{r clean-all}
all$OfficerRace <- gsub("H/L", "L", all$OfficerRace)
all$OfficerRace <- gsub("H", "L", all$OfficerRace)
all$OfficerRace <- gsub("Unknown", "U", all$OfficerRace)
all$OfficerRace <- gsub("m/m", "U", all$OfficerRace)
all$OfficerRace <- gsub("WHITE", "W", all$OfficerRace)
all$OfficerRace <- gsub("WLITE", "W", all$OfficerRace)
all$OfficerRace <- gsub("BLACK", "B", all$OfficerRace)
all$OfficerRace <- gsub("ASIAN", "A", all$OfficerRace)
all$OfficerRace <- gsub("Other", "O", all$OfficerRace)
all$OfficerRace <- gsub("AI/AN", "O", all$OfficerRace)
all$OfficerRace <- gsub("A/PI ", "A;", all$OfficerRace)
all$OfficerRace <- gsub("A/PI", "A", all$OfficerRace)
all$OfficerRace <- gsub("Multi-Racial", "O", all$OfficerRace)
all$OfficerRace <- gsub(",", ";", all$OfficerRace)
all$OfficerRace <- gsub(" ", "", all$OfficerRace)
all$OfficerRace <- gsub("A/W", "O", all$OfficerRace)
all$OfficerRace <- gsub("NA/W", "O", all$OfficerRace)
all$OfficerRace <- gsub("W/L", "O", all$OfficerRace)
all$OfficerRace <- gsub("W/A", "O", all$OfficerRace)

all <- all %>% 
  select(-c("Date", "NatureOfStop", "Notes", "SubjectAge", "NumberOfShots", "Department", "FullNarrative", "fdate", "month", "day", "tr", "AvgAge", "VA", "F", "nshots", "sit")) %>% 
  mutate(b = as.numeric(b > 0),
         w = as.numeric(w > 0),
         l = as.numeric(l > 0),
         a = as.numeric(a > 0),
         other = as.numeric(other), 
         replica = as.numeric(replica), 
         knife = as.numeric(knife), 
         gun = as.numeric(weapon == "gun"), 
         unarmed = as.numeric(weapon == "unarmed"),
         unknown_weapon = as.numeric(weapon == "U"),
         id = paste0("C", row_number())) %>% 
  select(-weapon)

all <- rename(all, sb = b, sl = l, sw = w, sa = a)

all$ob <- as.numeric(str_count(all$OfficerRace, "B") > 0)
all$ow <- as.numeric(str_count(all$OfficerRace, "W") > 0)
all$ol <- as.numeric(str_count(all$OfficerRace, "L") > 0)
all$oa <- as.numeric(str_count(all$OfficerRace, "A") > 0)

@@

|        As a result, I have decided to build off of VICE News’ study and investigate data on police shootings further to understand the roles that the race of both victim and officer, as well as other factors such as their genders, whether they are carrying a weapon, and the total number of victims in the crime, play in fatal versus non-fatal police shootings. Additionally, to account for the varying locations of the homicides, I will be using a dataset found on Kaggle detailing the gun provisions that are upheld by each state see how gun legislation affects lethal vs. non-lethal shootings. I will also add a predictor indicating whether the state in which the homicide occurred requires de-escalation training for police officers. Finally, I will explore how the ratio of black to white population in each location relates to police use of deadly force. This will allow me to better understand how racism has manifested itself in America’s police system and determine whether current attempts to prevent police use of lethal force are effective or not.

@@

|       Finally, to examine city demographics as a predictor, I will be using population data from the 2013 American Community Survey that VICE News had already cleaned and standardized to the shootings dataset, meaning the city names can be matched up between the two datasets. VICE News most likely provided only 2013 census data because it is the average year of all the years represented in the shootings dataset. The census data includes 7 variables: the police department and the city it's located in, as well as the city's Black, Asian, Hispanic, White, and overall total population in 2013.

@@

```{r}
single_victims <- all %>% 
  filter(!str_detect(SubjectRace, ";"))
nrow(single_victims)

mult_victims <- all %>% 
  filter(str_detect(SubjectRace, ";"))
nrow(mult_victims)

unknown <- all %>% 
  filter(str_detect(SubjectRace, "U"))
```
|       Upon first glance, there appears to be 3906 incidents in which there was only one victim, and 210 with multiple victims. This means that there is one incident in which the victim race was marked as "NA." Additionally,

```{r}
single_victims <- all %>% 
  filter(!str_detect(SubjectRace, ";"))

mult_victims <- all %>% 
  filter(str_detect(SubjectRace, ";"))

ggplot(single_victims, aes(x = Fatal, y = count)) +
  geom_bar(stat="identity") + 
  labs(x = "Fatal", y = "Count", title = "Counts for Fatal vs. Non-Fatal Police Shootings") 
  group_by(Fatal, race) %>% 
  summarise(count = n())
```
